### 2022-12-24 tex/lnd.tex
[ edit ] another draft of chapters 5 and 6

%This should stay fairly general, so it could be compiled to a DILL-style
%calculus as well as the one with usage annotations.
%Our framework compiles a ND system down to a sequent calculus.
%It isolates the building blocks of typing rules ($\vdash$, $\wedge$, $*$,
%$\Box$, $\cdot$) so that we can do more generic programming (e.g., usage
%checker).

@@

  Connectives:
  \vspace{-1em}
  \begin{multicols}{2}
    \noindent\ExecuteMetaData[\Bunchedtex]{PointwiseUnit}
    \noindent\ExecuteMetaData[\Bunchedtex]{forallb}
    \columnbreak
    \AgdaNoSpaceAroundCode{}
    \noindent\ExecuteMetaData[\Bunchedtex]{PointwiseConjunction}
    \vspace{-0.25em}
    \noindent\ExecuteMetaData[\Bunchedtex]{PointwiseImplication}
    \AgdaSpaceAroundCode{}
  \end{multicols}
  \vspace{-2em}
  \ExecuteMetaData[\Bunchedtex]{BunchedUnit}
  \ExecuteMetaData[\Bunchedtex]{BunchedConjunction}
  \ExecuteMetaData[\Bunchedtex]{BunchedImplication}
  \ExecuteMetaData[\Bunchedtex]{BunchedScaling}
  \caption{The bunched connectives}
  \label{fig:bunched}
\end{figure}

@@

The first three connectives are those we've already seen for intuitionistic
systems --- $\dot1$, $\dottimes$, and $\dotto$.
The absence of premises is encoded by $\dot1$, while the space between premises
sharing a context is encoded by $\dottimes$.
When we interpret a typing rule as a constructor of an inductive definition,
$\dotto$
interprets the horizontal line, reflecting the fact that the usage annotations
we start off with in the premises are those of the conclusion, corresponding to
a general principle of resource conservation.
The prototypical rules that use $\dot1$ and $\dottimes$ are the introduction
rules for $\top$ and $\with$, respectively.

@@

The full system \name{} is stated in terms of bunched connectives in
\cref{fig:lr-bunched}.
The bunched connectives also yield a reasonably consise definition of the Agda
data type of \name{} derivations, as seen in \cref{fig:lr-bunched-Agda}.

While we have seen a connection between the bunched connectives and the
connectives of $\name$, the two should not be confused.
In particular, the bunched connectives obey different laws to what we would
expect from linear logic.
For example, it would make sense to define a bunched connective $\dotplus$,
defined analogously to $\dottimes$.
This $\dotplus$ could be used to rephrase the introduction rules for $\oplus$.
We then have maps both ways between $T \dottimes (U \dotplus V)$ and
$(T \dottimes U) \dotplus (T \dottimes V)$, reminiscent of the distributivity of
additive connectives in bunched logic,
whereas linear logic only has a map from $(A \with B) \oplus (A \with C)$ to
$A \with (B \oplus C)$, and not a map the other way.
Looking at the interpretations, the connection with bunched logic makes sense.
Instead of the partial commutative monoid (often representing heaps) found in
standard semantics of bunched logic, we have a left semimodule of usage
contexts, which we are similarly interested in splitting and sharing between
various subterms.

A discrepancy with bunched logic is that $\dotto$ is not part of the Cartesian
closure of context-indexed families.
In fact, it fails to be a functor because it does not respect the ordering of
usage contexts.
This fact has little impact on the development of this thesis, because I mainly
use $\dotto$ at the top level, directly in conjunction with $\forallb{-}$.
The combination describes the ``external hom'', i.e.\ the morphisms in the
category of context-indexed families, with the internal hom being more like the
Kripke function space, which I do not use.

%\todo{Split chapter here. Play around with other systems using bunched
%  connectives. Lists/recursion, annotated arrow, Granule elim, sharing elim.}

@@

>>>>

@@

  Connectives:
  \vspace{-1em}
  \begin{multicols}{2}
    \noindent\ExecuteMetaData[\Bunchedtex]{PointwiseUnit}
    \noindent\ExecuteMetaData[\Bunchedtex]{PointwiseConjunction}
    \columnbreak
    %\AgdaNoSpaceAroundCode{}
    \noindent\ExecuteMetaData[\Bunchedtex]{Entails}
    %\AgdaSpaceAroundCode{}
  \end{multicols}
  \vspace{-2em}
  \ExecuteMetaData[\Bunchedtex]{BunchedUnit}
  \ExecuteMetaData[\Bunchedtex]{BunchedConjunction}
  \ExecuteMetaData[\Bunchedtex]{BunchedImplication}
  \ExecuteMetaData[\Bunchedtex]{BunchedScaling}
  \caption{The bunched connectives}
  \label{fig:bunched}
\end{figure}

@@

The first two connectives are those we've already seen for intuitionistic
systems --- $\dot1$ and $\dottimes$.
The absence of premises is encoded by $\dot1$, while the space between premises
sharing a context is encoded by $\dottimes$.
As for implication, I temporarily avoid giving a $\dotto$ connectives, instead
fusing it together with $\forallb{-}$ to produce the \emph{set} of
context-preserving functions $T \rightarrowtriangle U$.
When we interpret a typing rule as a constructor of an inductive definition,
$\rightarrowtriangle$
interprets the horizontal line, reflecting the fact that the usage annotations
we start off with in the premises are those of the conclusion, corresponding to
a general principle of resource conservation.
The prototypical rules that use $\dot1$ and $\dottimes$ are the introduction
rules for $\top$ and $\with$, respectively.

@@

\subsection{$\name$ stated using bunched connectives}\label{sec:lr-bunched}
The full system \name{} is stated in terms of bunched connectives in
\cref{fig:lr-bunched}.
The bunched connectives also yield a reasonably consise definition of the Agda
data type of \name{} derivations, as seen in \cref{fig:lr-bunched-Agda}.

@@

\subsection{Connection with bunched logic}\label{sec:bunched-logic}
While we have seen a connection between the bunched connectives and the
connectives of $\name$, the two should not be confused.
In particular, the bunched connectives obey different laws to what we would
expect from linear logic.
For example, it would make sense to define a bunched connective $\dotplus$,
defined analogously to $\dottimes$.
This $\dotplus$ could be used to rephrase the introduction rules for $\oplus$.
We then have maps both ways between $T \dottimes (U \dotplus V)$ and
$(T \dottimes U) \dotplus (T \dottimes V)$, reminiscent of the distributivity of
additive connectives in bunched logic,
whereas linear logic only has a map from $(A \with B) \oplus (A \with C)$ to
$A \with (B \oplus C)$, and not a map the other way.
Looking at the interpretations, the connection with bunched logic makes sense.
Instead of the partial commutative monoid (often representing heaps) found in
standard semantics of bunched logic, we have a left semimodule of usage
contexts, which we are similarly interested in splitting and sharing between
various subterms.

From bunched logic, we would expect the Cartesian product $\dottimes$ to have an
internal hom.
In the intuitionistic case, $\dotto$ filled this role.
However, with usage contexts, it makes sense for open types to be presheaves
over the partial order of contexts under pointwise $\leq$ of usage annotations.
The family $T \dotto U$ does not satisfy the functoriality condition because of
the contravariance in the domain $T$.
Instead, as found in models of bunched logic, we would want a Kripke function
space, like
$\lambda\Gamma.~\forall \Gamma' \leq \Gamma.~T\,\Gamma' \to U\,\Gamma'$.
However, I do not make use of such a connective.

The separating conjunction $\sep$ can be seen as a decategorified version of Day
convolution~\citep{Day70}.
It also resembles the use of ternary frames \todo{cite} in semantics of
substructural logics.

\subsection{Operations on bunched connectives}\label{sec:bunched-op}
To manipulate terms and other open types defined using bunched connectives, we
need the zero, addition, and multiplication relations to satisfy some laws.
For example, to achieve a symmetry map $T \sep U \rightarrowtriangle U \sep T$,
we need addition to satisfy the commutativity law
$\forall x,y,z : A.~x \leq y + z \to x \leq z + y$.

For all uses of bunched connectives in this thesis, the carrier set $A$ will
form a partial order --- for example, with contexts, the order is given by the
pointwise order on the usage vectors.
We then consider the category whose objects are posets and whose morphisms are
relations $R : A \rel B$ satisfying the contravariant-covariant law
$\forall x,x',y,y'.~x' \leq x \to y \leq y' \to xRy \to x'Ry'$.
This category can be given the usual monoidal product of relations, which is
the pointwise product of posets on objects.
Then, we will always expect zero and addition to together form a cocommutative
comonoid in this category.
With this structure, we can get the following equivalences and functions.

\begin{mathpar}
  I^* \sep T \leftrightarrowtriangle T \and
  T \sep I^* \leftrightarrowtriangle T \and
  (T \sep U) \sep V \leftrightarrowtriangle T \sep (U \sep V) \and
  T \sep U \leftrightarrowtriangle U \sep T \and
  (T \wand U) \sep T \rightarrowtriangle U \and
  I^* \wand T \leftrightarrowtriangle T \and
  (T \sep U) \wand V \leftrightarrowtriangle T \wand (U \wand V)
\end{mathpar}

I do not use algebraic properties of multiplication in conjunction with
manipulation of bunched connectives in this thesis, but we could expect
scalar multiplication
to add a comodule structure over cosemiring $R$ to the cocommutative comonoid
given by zero and addition.

### 2022-12-24 tex/lr.tex
[ edit ] another draft of chapters 5 and 6

  This posemiring can be used for sensitivity analysis, where we want to bound
  the effect of a perturbation on inputs in terms of some semantic notion of
  distance between values.
  An annotation $\gr r$ says that if that input is perturbed by at most $r$,
  then the output will change by at most $1$.
  An $\gr\infty$-annotated variable gives the very strong guarantee that any
  change in the input will make a minimal change to the output, while a
  $\gr0$-annotated variable provides no guarantee at all.
  Addition forbids general contraction, which would otherwise allow arbitrary
  finite blow-up of the effect of any non-$\gr0$-annotated variable.
  However, the ordering, with $\gr0$ at the top, means that we have general
  weakening, so the resulting sensitivity calculus has an affine flavour.
\end{example}

>>>>

  The intended semantics for these annotations is to interpret types as metric
  spaces and terms as Lipschitz-continuous maps.
  That is to say, each type $A$ comes with a notion of distance $d_A$, and a map
  $f : A \to B$ satisfies
  \[
    \exists\gr r \in \mathbb R^+.\forall x,y.~%
    d_B(f(x), f(y)) \leq \gr rd_A(x, y).
  \]
  I internalise this constant $\gr r$ so that derivations of
  $\gr rA \vdash B$ are interpreted as $\gr r$-Lipschitz-continuous functions.
  If $\gr r = \gr\infty$, then the Lipschitz condition is trivially satisfied,
  meaning that $\gr\infty$-annotated variables can be used without constraint.
  If $\gr r = \gr0$, then $f$ has to be constant, and the corresponding
  syntactic constraint is that a $\gr0$-annotated variable cannot be used.

  Addition forbids general contraction, which would otherwise allow arbitrary
  finite blow-up of the effect of any non-$\gr0$-annotated variable.
  However, the ordering, with $\gr0$ at the top, means that we have general
  weakening, so the resulting sensitivity calculus has an affine flavour.

  Such a system has been applied to sensitivity analysis, a component of
  differential privacy, by \citet{reed10distance}.
\end{example}

### 2022-12-24 tex/lrsub.tex
[ edit ] another draft of chapters 5 and 6

Given a \AgdaRecord{Kit}, \cref{thm:lr-bind} looks like the following.
The \AgdaField{lookup} clauses still contain essentially the same structure as
in the intuitionistic case: discriminating on whether the variable is old or
new, using the given environment \AgdaBound{$\rho$} and weakening on the old
variables, and using \AgdaField{vr} to repackage new variables.
I will not explain any of the algebraic manipulations here; see
\cref{thm:lr-bind}.

@@

Finally, the simultaneous substitution \AgdaFunction{sub} specialises to
single substitution \AgdaFunction{sub[-]}.
Single substitution is stated as an admissible rule below.
To substitute in for $\gr r$-many $A$ in the second term, we need to derive
one $A$ with usages $\grP$, and then assert that the result can handle the
usages of the original term $\grQ$, plus $\gr r$-many copies of $\grP$.

\[
  \ebrule{%
    \hypo{\grR \leq \grQ + \gr r\grP}
    \hypo{\grP\gamma \vdash A}
    \hypo{\grQ\gamma, \gr rA \vdash B}
    \infer3[singleSub]{\grR\gamma \vdash B}
  }
\]

The proof strategy for producing the substitution \AgdaFunction{$\sigma$} is
to proceed structurally on the codomain context $\grQ\gamma, \gr rA$ using
\cref{thm:construct-env}, applying the identity substitution
\AgdaFunction{id\^{}Env} on the $\gamma$ half, and dropping the term
\AgdaBound{t} in place of the variable we are substituting for.

\ExecuteMetaData[\LRKits]{subSingle}

>>>>

Given a \AgdaRecord{Kit}\AgdaSpace{}\AgdaBound{$\V$},
\cref{thm:lr-bind} gives a function with the following type.

@@

Finally, the simultaneous substitution \AgdaFunction{sub} specialises to
single substitution.

\begin{corollary}[Single substitution]\label{thm:single-sub}
  The following equivalent rules are admissible.
  \begin{mathpar}
    \ebrule{%
      \hypo{\grR \leq \gr r\grP + \grQ}
      \hypo{\grP\gamma \vdash A}
      \hypo{\grQ\gamma, \gr rA \vdash B}
      \infer3{\grR\gamma \vdash B}
    }
    \and
    \ebrule[comb]{%
      \hypo{\gr r \cdot \plr{{} \vdash A}}
      \hypo{\sep}
      \hypo{\gr rA \vdash B}
      \infer3{{} \vdash B}
    }
  \end{mathpar}
\end{corollary}
\begin{proof}
  It is enough to construct a substitution of type
  $\grR\gamma \env\vdash \grQ\gamma, \gr rA$.
  To do this, we use \cref{thm:construct-env} cases $\sep(\rightarrowtriangle)$
  and $\cdot(\rightarrowtriangle)$ on inequalities
  $\grR \leq \grQ + \gr r\grP$ and $\gr r\grP \leq \gr r\grP$ respectively to
  leave us needing a substitution of type $\grQ\gamma \env\vdash \grQ\gamma$ and
  a term of type $\grP\gamma \vdash A$.
  For the substitution, we give the identity substitution (\cref{thm:env-id}),
  and we have the term as a hypothesis.
\end{proof}

### 2022-12-24 tex/rec.tex
[ edit ] another draft of chapters 5 and 6

\subsection{Alternative object-language connectives}

@@

We can also give an alternative elimination rule for $\otimes$-products inspired
by Granule~\citep{Granule18,HMWO21}.
This modified rule allows us to, for example, have a variable
$\gr rx : A \otimes B$ for any $\gr r$, and pattern-match it into
$\gr ry : A, \gr rz : B$.
With the standard $\name$ rule from \cref{fig:lr}, the newly bound variables are
always given annotation $\gr 1$, and we would only be able to do the match in
the first place if $\gr r$ were coercible to (i.e.\ less than or equal to)
$\gr1$.
The alternative rule below is not compatible with linear logic, because it
allows us to derive
$\oc\gr r\plr{A \otimes B} \multimap \oc\gr rA \otimes \oc\gr r B$
parametric in $\gr r$, $A$, and $B$.

@@

  \begin{align*}
    \text{nil}\left[
    \begin{prooftree}
      \hypo{\grR \leq \gr0}
      \infer1[$I$-I]{\grR\gamma \vdash I}
      \infer1[$\oplus$-I$_0$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
    &&
    \text{cons}\left[
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma \vdash A}
      \hypo{\grQ\gamma \vdash \mathrm{List}_A}
      \infer3[$\otimes$-I]{\grR\gamma \vdash A \otimes \mathrm{List}_A}
      \infer1[$\oplus$-I$_1$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
  \end{align*}
  \begin{displaymath}
    \text{foldr}\left[
    \begin{prooftree}
      \hypo{\grP\gamma \vdash \mathrm{List}_A}
      \infer0[Var]{\gr0\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash I \oplus \plr{A \otimes C}}
      \hypo{\nabla^n}
      \hypo{\nabla^c}
      \infer3[$\oplus$-E]{\grQ\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash C}
      \infer2[$\mu$-E]{\grR\gamma \vdash C}
    \end{prooftree}
    \right.
  \end{displaymath}
  \begin{align*}
    \textrm{where }\nabla^n &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1I \vdash I}
      \hypo{\grQ\gamma \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0I \vdash C}
      \infer2[$I$-E]{\grQ\gamma, \gr1I \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1I
        \vdash C}
    \end{prooftree}
    \\\\
    \textrm{and }\nabla^c &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1\plr{A \otimes C} \vdash A \otimes C}
      \hypo{\grQ\gamma, \gr1A, \gr1C \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{A \otimes C}, \gr1A, \gr1C \vdash C}
      \infer2[$\otimes$-E]{\grQ\gamma, \gr1\plr{A \otimes C} \vdash C}
      \infer1[Wk]%
      {\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1\plr{A \otimes C}
        \vdash C}
    \end{prooftree}
  \end{align*}
\end{example}

@@

@@

>>>>

\subsection{Alternative object-language connectives}\label{sec:alt}

@@

We can also give an alternative elimination rule for $\otimes$-products, as
found in Granule~\citep{Granule18,HMWO21} and the work of
\citet{AbelBernardy2020} and \citet{reed10distance}.
This modified rule allows us to, for example, have a variable
$\gr rx : A \otimes B$ for any $\gr r$, and pattern-match it into
$\gr ry : A, \gr rz : B$.
With the standard $\name$ rule from \cref{fig:lr}, the newly bound variables are
always given annotation $\gr 1$, and we would only be able to do the match in
the first place if $\gr r$ were coercible to (i.e.\ less than or equal to)
$\gr1$.
The alternative rule below is not compatible with linear logic, because it
allows us to derive
$\oc\gr r\plr{A \otimes B} \multimap \oc\gr rA \otimes \oc\gr r B$
parametric in $\gr r$, $A$, and $B$.

@@

  \begin{align*}
    \TirName{nil}\left[
    \begin{prooftree}
      \hypo{\grR \leq \gr0}
      \infer1[$I$-I]{\grR\gamma \vdash I}
      \infer1[$\oplus$-I$_0$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
    &&
    \TirName{cons}\left[
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma \vdash A}
      \hypo{\grQ\gamma \vdash \mathrm{List}_A}
      \infer3[$\otimes$-I]{\grR\gamma \vdash A \otimes \mathrm{List}_A}
      \infer1[$\oplus$-I$_1$]%
      {\grR\gamma \vdash I \oplus \plr{A \otimes \mathrm{List}_A}}
      \infer1[$\mu$-I]{\grR\gamma \vdash \mathrm{List}_A}
    \end{prooftree}
    \right.
  \end{align*}
  \begin{displaymath}
    \TirName{foldr}\left[
    \begin{prooftree}
      \hypo{\grP\gamma \vdash \mathrm{List}_A}
      \infer0[Var]{\gr0\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash I \oplus \plr{A \otimes C}}
      \hypo{\nabla^n}
      \hypo{\nabla^c}
      \infer3[$\oplus$-E]{\grQ\gamma, \gr1\plr{I \oplus \plr{A \otimes C}}
        \vdash C}
      \infer2[$\mu$-E]{\grR\gamma \vdash C}
    \end{prooftree}
    \right.
  \end{displaymath}
  \begin{align*}
    \textrm{where }\nabla^n &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1I \vdash I}
      \hypo{\grQ\gamma \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0I \vdash C}
      \infer2[$I$-E]{\grQ\gamma, \gr1I \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1I
        \vdash C}
    \end{prooftree}
    \\\\
    \textrm{and }\nabla^c &\coloneqq
    \begin{prooftree}
      \infer0[Var]{\gr0\gamma, \gr1\plr{A \otimes C} \vdash A \otimes C}
      \hypo{\grQ\gamma, \gr1A, \gr1C \vdash C}
      \infer1[Wk]{\grQ\gamma, \gr0\plr{A \otimes C}, \gr1A, \gr1C \vdash C}
      \infer2[$\otimes$-E]{\grQ\gamma, \gr1\plr{A \otimes C} \vdash C}
      \infer1[Wk]%
      {\grQ\gamma, \gr0\plr{I \oplus \plr{A \otimes C}}, \gr1\plr{A \otimes C}
        \vdash C}
    \end{prooftree}
  \end{align*}
\end{example}

@@

Among the inductive types given by this $\mu$ operator are also
\emph{$\with$-lists} (\emph{with-lists}).

\begin{example}
  Let us define \emph{$\with$-lists} as follows.
  \[
    \mathrm{WithList}_A \coloneqq \mu X.~\top \oplus \plr{A \with X}
  \]
  Unlike ordinary lists (which we may call \emph{$\otimes$-lists}),
  $\with$-lists support a \emph{lookup} operation with the type below, where
  $\mathbb N \coloneqq \mu X.~I \oplus X$.
  \begin{align*}
    \mathit{lookup} :
    \mathrm{WithList}_A \multimap \mathbb N \multimap \top \oplus A
  \end{align*}
  The $\name$ term implementing $\mathit{lookup}$ in terms of iterators
  (\TirName{$\mu$-E}) is complex and tedious, so I do not show it here.
  Instead, I provide the following clausal specification, where
  $\mathrm{nil}\plr{-}$ and $\mathrm{cons}\plr{-}$ are the constructors of
  $\with$-lists, $\mathrm{zero}\plr{}$ and $\mathrm{succ}\plr{}$ are the
  constructors of $\mathbb N$, $\mathrm{inl}\plr{-}$ and $\mathrm{inr}\plr{-}$
  are the constructors of $\oplus$, $\langle\rangle$ constructs the
  inhabitant of $\top$, and postfix $.\pi_l$ and $.\pi_r$ are the projections
  out of a $\with$-pair.
  %\begin{align*}
  %  \mathit{lookup}\,\mathrm{nil}\plr{t}\,i
  %  &\coloneqq \mathrm{inl}\plr{\langle\rangle} \\
  %  \mathit{lookup}\,\mathrm{cons}\plr{\mathit{xxs}}\,\mathrm{zero}\plr{}
  %  &\coloneqq \mathrm{inr}\plr{\mathit{xxs}.\pi_l} \\
  %  \mathit{lookup}\,\mathrm{cons}\plr{\mathit{xxs}}\,\mathrm{succ}\plr{i}
  %  &\coloneqq \mathit{lookup}\,\plr{\mathit{xxs}.\pi_r}\,i
  %\end{align*}

  %\noindent{}\hspace{-0.75ex}
  \[
    \begin{pmboxed}
      \defaultcolumn{l}
      \> \mathit{lookup} \> \mathrm{nil}\plr{t} \> i
      \> \coloneqq \> \mathrm{inl}\plr{\langle\rangle} \\
      \> \mathit{lookup} \> \mathrm{cons}\plr{\mathit{xxs}}
      \> \mathrm{zero}\plr{} \> \coloneqq
      \> \mathrm{inr}\plr{\mathit{xxs}.\pi_l} \\
      \> \mathit{lookup} \> \mathrm{cons}\plr{\mathit{xxs}}
      \> \mathrm{succ}\plr{i} \> \coloneqq
      \> \mathit{lookup}\,\plr{\mathit{xxs}.\pi_r}\,i \\
    \end{pmboxed}
  \]
  %\begin{verbatim}
  %  lookup := \xs. wfoldr(xs;
  %    List I -o T + A;
  %    t. \i. foldr(i; T -o T + A; \t'. inl(t'); u, r. let () := u in r) t;
  %    xr. \i. let (i', r) :=
  %      foldr(i;
  %        A & (T + A) -o List I * (T + A);
  %        \xr'. ([], prl(xr'));
  %        u, ir. \xr'.
  %          let () := u in
  %          let (i', r) := ir in
  %          (() :: i', prr(xr') prl(ir)))
  %      xr in
  %      foldr(i'; T + A; r; u, r'. let () := u in r'))
  %\end{verbatim}
  %\begin{align*}
  %  \mathit{lookup} &:
  %  \mathrm{WithList}_A \multimap \mathrm{List}_I \multimap \top \oplus A
  %  \\
  %                  &\coloneqq \lambda\mathit{xs}.~\mathrm{rec}\plr{xs; %
  %                    \mathrm{List}_I \multimap \top \oplus A; %
  %                    t.~\lambda i.~\mathrm{inl}\plr{%
  %                    \mathrm{foldr}\plr{i, \top \multimap \top; \lambda t'.~t'; %
  %                    u, r.~\textrm{let }{*} \coloneqq u\textrm{ in }r}\,t}; %
  %                    \mathit{xr}.~\lambda i.~%
  %                    \mathrm{foldr}\plr{i; %
  %                    A \with \plr{\top \oplus A} \multimap \top \oplus A; %
  %                    \lambda
  %                    }\,\mathit{xr}}
  %\end{align*}
\end{example}

@@

A similar modality, $\Box$, appears in the work of \citet{BB18} when they deal
with \emph{persistent predicates} in bunched logic.
They similarly aim to pick out those objects not depending on any external
resources.
